---
title: "Darwin Core Mapping"
subtitle: "For: Florient - Wild plant occurrences"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # date transformation
##library(ggmap)          # geocoding data
library(tidygeocoder)   # geocoding
library(mapdata)
#library(BelgiumMaps.StatBel)

```

# Read source data

Create a data frame `input_data` from the source data:
The source data was corrected in Excel
Muskrat occurrences opened in openRefine
Obsolete columns removed
some columns renamed to DwC term
File exported to csv

```{r}
#input_interim <- read_csv2(file = here:here("datasets", "florient-occurrences", "data", "raw", "florient.csv"))  ",locale(encoding = "UTF-8")" --> no good?

input_interim <- read_csv2(file = here::here("datasets", "florient-occurrences", "data", "raw", "florient.csv"))
```

Preview data:

```{r}
input_interim %>% head(n = 5)
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_interim %<>% remove_empty("rows")
```

```{r}
input_interim %>% head(n = 5)
```

# Georeference source data



## geocoding the source data

```{r}
input_interim %<>% mutate(country = "Belgium")
```

```{r eval=FALSE, include=FALSE}
input_interim %<>% geocode(city = deelgemeente_of_dorp , country = country)

```

## save geocoded file

```{r eval=FALSE, include=FALSE}
write_csv(input_interim, here::here("datasets/florient-occurrences", "data", "interim", "input_geocoded2.csv"), na = "")
```


# Occurrence core

```{r eval=FALSE, include=FALSE}
input_geocoded <- read_csv(file = here::here("datasets", "florient-occurrences", "data", "interim", "input_geocoded.csv"))
```


## Pre-processing

```{r}
occurrence <- input_geocoded
```



## Remove obsolete columns

```{r}
occurrence %<>% select(-c(arrondissement_Van_Hoorebeke,hoofdgemeente))
```

# Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

# Event

### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "Ghent University") # e.g. "INBO"
```

### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "https://toComplete")
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "UGent") # e.g. "INBO"
```

### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Florient - a database on wild vascular plants in East Flanders in the beginning of the 19th century") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

The following terms contain information about the taxon:

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### informationWithHeld

### dataGeneralizations

### occurrenceID

**This need to be fixed**

```{r}
occurrence %<>% mutate(row_id = row_number()) %>%
                mutate(dwc_occurrenceID = str_c("UGENT:FLORIENT:OCC:", row_id))
   
```

### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = 'Charles Van Hoorebeke')
```

### individualCount


### organismQuantity

### organismQuentityType

### sex


### lifeStage


### behavior

### occurrenceRemarks

```{r}
# occurrence %<>% mutate(occurrenceRemarks = 'action_en')
```

### samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingProtocol = 'plant collecion')
```

### samplingEffort

### eventDate

```{r}
occurrence %<>% mutate(dwc_eventDate = str_replace_all(periode, "-","/")) 
```

# Location

```{r}
occurrence %<>% mutate(dwc_decimalLongitude = long) %>%
                mutate(dwc_decimalLatitude = lat) %>%
                mutate(dwc_geodeticDatum = "WGS84") 
```

```{r eval=FALSE, include=FALSE}
head(occurrence, n = 5)
occurrence %<>%
   mutate(dwc_decimalLongitude = round(decimalLongitude, digits = 5)) %>%
   mutate(dwc_decimalLatitude = round(decimalLatitude, digits = 5))
```

```{r}
occurrence %<>%   
   mutate(dwc_decimalLatitude = as.character(format(dwc_decimalLatitude, nsmall = 5))) %>%
   mutate(dwc_decimalLongitude = as.character(format(dwc_decimalLongitude, nsmall = 5))) 
   
```

### continent

```{r}
occurrence %<>% mutate(dwc_continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
occurrence %<>% mutate(dwc_countryCode = "BE") # e.g. "Belgium = BE"
```

### municipality

municipality already in source file

```{r}
occurrence %<>%
  mutate(dwc_municipality = deelgemeente_of_dorp)
```

### locationID


### verbatimCoordinates

### verbatimLatitude

### verbatimLongitude

### verbatimCoordinateSystem


### verbatimSRS


### decimalLatitude

### decimalLongitude

### geodeticDatum

### coordinateUncertaintyInMeters

```{r}
occurrence %<>% mutate(dwc_coordinateUncertaintyInMeters = '5000' )
```


### georeferenceRemarks

```{r}
occurrence %<>% mutate(dwc_georeferenceRemarks = 'coordinate is generated based on municipality by tidyGeocoder in R' )
```

### identifiedBy

```{r}
occurrence %<>% mutate(dwc_identifiedBy = bron)
```

### scientificName 

```{r}
head(occurrence, n=5)
occurrence %<>% mutate(dwc_scientificName = Wetenschappelijke_naam)
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Plantae")
```

### scientificNameAuthorship


### verbatimScientificName

```{r}
occurrence %<>% mutate(dwc_verbatimScientificName = Nederlandse_naam)
```

# taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = "species")
```

# nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICBN") # e.g. "ICZN"
```

### occurrenceStatus

```{r}
occurrence %<>% mutate(dwc_occurrenceStatus = "present") 
```

## Plotting data on the map





## Post-processing

```{r}

occurrence %<>% select(starts_with("dwc_"))  


```

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
```

```{r}
head(occurrence, n = (5))
```
#visualize data



Define the order of the output columns

```{r eval=FALSE, include=FALSE}
col_order <- c( "type","language","license","rightsHolder","accessRights","datasetID"
               ,"institutionCode","datasetName","basisOfRecord","occurrenceID","recordedBy"
               ,"individualCount","occurrenceStatus","eventDate","locationID", "continent","countryCode"
               ,"verbatimLatitude","verbatimLongitude","verbatimCoordinateSystem","verbatimSRS"
               ,"decimalLatitude","decimalLongitude","geodeticDatum","coordinateUncertaintyInMeters"
               ,"identifiedBy","scientificName","kingdom","scientificNameAuthorship","taxonRank","nomenclaturalCode"
               ) 
               # still needed for this dataset 
               # removed for this dataset,"occurrenceRemarks","municipality","samplingProtocol","samplingEffort"
occurrence <- occurrence[, col_order]
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets/florient-occurrences", "data", "processed", "occurrence.csv"), na = "")



```
