---
title: "Darwin Core Mapping"
subtitle: "For: Florient - Wild plant occurrences"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # date transformation
##library(ggmap)          # geocoding data
library(tidygeocoder)   # geocoding
library(mapdata)
library(BelgiumMaps.StatBel)
```

# Read source data

Create a data frame `input_data` from the source data:
The source data was corrected in Excel
Muskrat occurrences opened in openRefine
Obsolete columns removed
some columns renamed to DwC term
File exported to csv

```{r}
#input_interim <- read_csv2(file = here:here("datasets", "florient-occurrences", "data", "raw", "florient.csv"))  

input_interim <- read_csv2(file = here::here("datasets", "florient-occurrences", "data", "raw", "florient.csv"))
```

Preview data:

```{r}
input_interim %>% head(n = 5)
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_interim %<>% remove_empty("rows")
```

```{r}
input_interim %>% head(n = 5)
```

# Georeference source data

## geocoding the source data
```{r}
input_interim %<>% geocode(deelgemeente_of_dorp)

```

# Occurrence core

## Pre-processing

Create a dataframe occurrence data only 

```{r}
occurrence <- input_interim
```

## Remove obsolete columns

```{r}
occurrence %<>% select(-c(arrondissement_Van_Hoorebeke,hoofdgemeente))
```

# Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

# Event

### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "Ghent University") # e.g. "INBO"
```

### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "https://toComplete")
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "UGent") # e.g. "INBO"
```

### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Florient - a database on wild vascular plants in East Flanders in the beginning of the 19th century") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

The following terms contain information about the taxon:

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### informationWithHeld

### dataGeneralizations

### occurrenceID

**This need to be fixed**

```{r}
occurrence %<>% mutate(row_id = row_number()) %>%
                mutate(dwc_occurrenceID = str_c("UGENT:FLORIENT:OCC:", row_id))
   
```

### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = bron)
```

### individualCount


### organismQuantity

### organismQuentityType

### sex


### lifeStage


### behavior

### occurrenceRemarks

```{r}
# occurrence %<>% rename(occurrenceRemarks = 'action_en')
```

### samplingProtocol

### samplingEffort

### eventDate

```{r}
occurrence %<>% mutate(dwc_eventDate = (periode)) 
```

# Location

```{r}
occurrence %<>% mutate(decimalLongitude = long) %>%
                mutate(decimalLatitude = lat) %>%
                mutate(dwc_geodeticDatum = "WGS84") 
```

```{r}
head(occurrence, n = 5)
occurrence %<>%
   mutate(decimalLongitude = round(decimalLongitude, digits = 5)) %>%
   mutate(decimalLatitude = round(decimalLatitude, digits = 5))
```

```{r}
occurrence %<>%   
   mutate(decimalLatitude = as.character(format(decimalLatitude, nsmall = 5))) %>%
   mutate(decimalLongitude = as.character(format(decimalLongitude, nsmall = 5))) 
   
```

### continent

```{r}
occurrence %<>% mutate(dwc_continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
occurrence %<>% mutate(dwc_countryCode = "BE") # e.g. "Belgium = BE"
```

### municipality

municipality already in source file

```{r}
occurrence %<>%
  mutate(dwc_municipality = deelgemeente_of_dorp)
```

### locationID


### verbatimCoordinates

### verbatimLatitude

### verbatimLongitude

### verbatimCoordinateSystem


### verbatimSRS


### decimalLatitude

### decimalLongitude

### geodeticDatum

### coordinateUncertaintyInMeters

### georeferenceRemarks

### identifiedBy

```{r}
occurrence %<>% mutate(dwc_identifiedBy = bron)
```

### scientificName 

```{r}
head(occurrence, n=5)
occurrence %<>% mutate(dwc_scientificName = Wetenschappelijke_naam)
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Plantae")
```

### scientificNameAuthorship


### verbatimScientificName

```{r}
occurrence %<>% mutate(dwc_verbatimScientificName = Nederlandse_naam)
```

# taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = "species")
```

# nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICBN") # e.g. "ICZN"
```

### occurrenceStatus

## Plotting data on the map

```{r}
# Personalised ggplot2 theme
theme_marine <- function(){
  theme_bw() +
    theme(axis.text = element_text(size = 16),
          axis.title = element_text(size = 20),
          axis.line.x = element_line(color="black"),
          axis.line.y = element_line(color="black"),
          panel.border = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.y = element_blank(),
          plot.margin = unit(c(1, 1, 1, 1), units = , "cm"),
          plot.title = element_text(size = 20),
          legend.text = element_text(size = 12),
          legend.title = element_blank(),
          legend.position = c(0.9, 0.9),
          legend.key = element_blank(),
          legend.background = element_rect(color = "black",
                                           fill = "transparent",
                                           size = 2, linetype="blank"))
}
```




## Post-processing

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")

```

Define the order of the output columns

```{r}
col_order <- c( "type","language","license","rightsHolder","accessRights","datasetID"
               ,"institutionCode","datasetName","basisOfRecord","occurrenceID","recordedBy"
               ,"individualCount","occurrenceStatus","eventDate","locationID", "continent","countryCode"
               ,"verbatimLatitude","verbatimLongitude","verbatimCoordinateSystem","verbatimSRS"
               ,"decimalLatitude","decimalLongitude","geodeticDatum","coordinateUncertaintyInMeters"
               ,"identifiedBy","scientificName","kingdom","scientificNameAuthorship","taxonRank","nomenclaturalCode"
               ) 
               # still needed for this dataset 
               # removed for this dataset,"occurrenceRemarks","municipality","samplingProtocol","samplingEffort"
occurrence <- occurrence[, col_order]
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets/mica-uvw-occurrences", "data", "processed", "occurrence.csv"), na = "")



```
